---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Tsion"
date: "02-08-2026"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

*Install dsbox first
```{r install-dsbox}
#manually downloaded and opened the two datasets.
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

```{r assign-shorter-values-for-datasets }
dn <- dennys
lq <- laquinta
```


### Exercise 1

```{r dennys-dimensions}
nrow(dn)
ncol(dn)
```

The Dennys dataset contains `r nrow(dn)` and `r ncol(dn)` columns. The rows represent a Denny's restaurant the data contains. The variables are address, city, state, zip (code), longitude, and latitude. 

### Exercise 2

```{r laquinta-dimensions}
nrow(lq)
ncol(lq)
```

The Laquinta dataset contains `r nrow(lq)` and `r ncol(lq)` columns. The rows represent a Laquinta hotel the data contains. The variables are address, city, state, zip (code), longitude, and latitude (same as the dennys' dataset). 

### Exercise 3

...

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...

 
Add exercise headings as needed.

