---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Tsion"
date: "02-08-2026"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

*Install dsbox first
```{r install-dsbox}
#manually downloaded and opened the two datasets.
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

```{r assign-shorter-values-for-datasets }
dn <- dennys
lq <- laquinta
```


### Exercise 1

```{r dennys-dimensions}
nrow(dn)
ncol(dn)
```

The Dennys dataset contains `r nrow(dn)` and `r ncol(dn)` columns. The rows represent a Denny's restaurant the data contains. The variables are address, city, state, zip (code), longitude, and latitude. 

### Exercise 2

```{r laquinta-dimensions}
nrow(lq)
ncol(lq)
```

The Laquinta dataset contains `r nrow(lq)` and `r ncol(lq)` columns. The rows represent a Laquinta hotel the data contains. The variables are address, city, state, zip (code), longitude, and latitude (same as the dennys' dataset). 

### Exercise 3

Laquinta hotels have locations in the following places outisde of US:

North America: Canada, Mexico

Asia: China

Australia & the Pacific Rim: New Zealnd

Europe: Georgia, Turkiye

Middle East: UAE (Abu Dhabi, Dubai)

South America: Colomibia, Ecuador

For Denny's, the link only shows the U.S locations. It lists "View U.S Locations (1257)" in the website. But our dataset contains 1643 observations so I suspected there might be locations outside the U.S that we just can't access through the website. But I am also not sure why the numbers in our dataset are higher since it looks like it's listing the (US) states and zip codes there too. 

I did a general google search, and wikipedia says, "It operates over 1,400 restaurants in the United States, Canada, México, Puerto Rico, and several other international locations." (https://en.wikipedia.org/wiki/Denny%27s) 

However, I wasn't able to find a specific link that shows the different Denny's locations outside the U.S. 

### Exercise 4



### Exercise 5



### Exercise 6



 
Add exercise headings as needed.

